\section{Introduction}
\subsection{Why \Lang{}? -- The Core Concept}
We propose to implement an elegant yet secure general purpose object-oriented programming language. Interesting features have been selected from the history of object-oriented programming and will be combined with the familiar ideas and style of modern languages.

\Lang{} combines three disparate but equally important tenants:


\begin{enumerate}
\item{Purely object-oriented 
    
    \Lang{} brings to the table a purely object oriented programming language where every type is
    modeled as an object--including the standard primitives. Integers, Strings, Arrays, and other types may be expressed in the standard fashion but are objects behind the scenes and can be treated as such.}

\item{Controllable

   \Lang{} provides innate security by choosing object level access
   control as opposed to class level access specifiers. Private members of one object
   are inaccessible to other objects of the same type. Overloading is not allowed.
   No subclass can turn your functionality on its head.}

\item{Versatile

    \Lang{} allows programmers to place "refinement methods" inside their code.
    Alone these methods do nothing, but may be defined by subclasses so as to extend
    functionality at certain important positions. Anonymous instantiation allows for
    extension of your classes in a quick easy fashion.}
\end{enumerate}

\subsection{ The Motivation Behind \Lang{}}
\Lang{} is a reaction to the object-oriented languages before it.
Obtuse syntax, flaws in security, and awkward implementations plague
the average object-oriented language. \Lang{} is intended as a step
toward ease and comfort as an object-oriented programmer.


The first goal is to make an object-oriented language that is comfortable
in its own skin. It should naturally lend itself to constructing API-layers
and abstracting general models. It should serve the programmer towards their
goal instead of exerting unnecessary effort through verbosity and awkwardness
of structure.


The second goal is to make a language that is stable and controllable.
The programmer in the lowest abstraction layer has control over how those
higher may procede. Unexpected runtime behavior should be reduced through
firmness of semantic structure and debugging should be a straight-forward
process due to pure object and method nature of \Lang{}.

\subsection{\Lang{} Feature Set}

\Lang{} will provide the following features:

\begin{itemize}
\item Universal objecthood
\item Optional ``refinement'' functions to extend superclass functionality
\item Anonymous class instantiation
\item Static typing
\item Access specifiers that respect object boundaries, not class boundaries
\end{itemize}

\subsection{\Compiler{}: The \Lang{} Compiler}

The compiler will proceed in two steps. First, the compiler will interpret
the source containing possible syntactic shorthand into a file
consisting only of the most concise and structurally sound GAMMA core. After this the compiler will transform
general patterns into (hopefully portable) C code, and compile this to
machine code with whatever compiler the user specifies.
