\subsection{Lexical Elements}
\subsubsection{Whitespace}
The new line (line feed), form feed, carriage return, and vertical tab characters will all be treated equivalently as vertical whitespace. Tokens are separated by horizontal (space, tab) and vertical (see previous remark) whitespace of any length (including zero).

\subsubsection{Identifiers}
Identifiers are used for the identification of variables,  methods and types. An identifer is a sequence of alphanumeric characters, uppercase and lowercase, and underscores. A type identifier must start with an uppercase letter; all others must start with a lower case letter. Additionally, the lexeme of a left bracket followed immediately by a right bracket -- \verb![]! -- may appear at the end of a type identifier in certain contexts, and that there may be multiple present in this case (denoting arrays, etc). The legal contexts for such will be described later.

\subsubsection{Keywords}
The following words are reserved keywords. They may not be used as identifiers:
\begin{center}
\begin{tabular}{cccccc}
\verb!and! & \verb!class! & \verb!else! & \verb!elsif! & \verb!extends! & \verb!false!\\
\verb!if! & \verb!init! & \verb!main! & \verb!nand! & \verb!new! & \verb!nor!\\
\verb!not! & \verb!or! & \verb!private! & \verb!protected! & \verb!public! & \verb!refinable!\\
\verb!refine! & \verb!refinement! & \verb!return! & \verb!super! & \verb!this! & \verb!to! \\
\verb!true! & \verb!void! & \verb!while! & \verb!xor!\\
\end{tabular}
\end{center}

\subsubsection{Operators}
There are a large number of (mostly binary) operators:
\begin{center}
\begin{tabular}{ccccccc}
\verb!=! & \verb!=/=! & \verb|<>| & \verb!<! & \verb!<=! & \verb!>! & \verb!>=!\\
\verb!+! & \verb!-! & \verb!*! & \verb!/! & \verb!%! & \verb!^! & \verb!:=!\\
\verb!+=! & \verb!-=! & \verb!*=! & \verb!/=! & \verb!%=! & \verb!^=! \\
\verb!and! & \verb!or! & \verb!not! & \verb!nand! & \verb!nor! & \verb!xor!&\verb!refinable!\\
\end{tabular}
\end{center}

\subsubsection{Literal Classes}
A literal class is a value that may be expressed in code without the use of the new keyword. These are the fundamental units of program.

\paragraph{Integer Literals}
An integer literal is a sequence of digits. It may be prefaced by a unary minus symbol. For example:
\begin{itemize}
\item \verb!777!
\item \verb!42!
\item \verb!2!
\item \verb!-999!
\item \verb!0001!
\end{itemize}

\paragraph{Float Literals}
A float literal is a sequence of digits and exactly one decimal point/period. It must have at least one digit before the decimal point and at least one digit after the decimal point. It may also be prefaced by a unary minus symbol. For example:
\begin{itemize}
\item \verb!1.0!
\item \verb!-0.567!
\item \verb!10000.1!
\item \verb!00004.70000!
\item \verb!12345.6789!
\end{itemize}

\paragraph{Boolean Literals}
A boolean literal is a single keyword, either \verb!true! or \verb!false!.

\paragraph{String Literals}
A string literal consists of a sequence of characters enclosed in double quotes. Note that a string literal can have the new line escape sequence within it (among others, see below), but cannot have a new line (line feed), form feed, carriage return, or vertical tab within it; nor can it have the end of file. Please note that the sequence may be of length zero. For example:
\begin{itemize}
\item \verb!"Yellow matter custard"!
\item \verb!""!
\item \verb!"Dripping\n   from a dead"!
\item \verb!"'s 3y3"!
\end{itemize}

The following are the escape sequences available within a string literal; a backslash followed by a character outside of those below is an error.
\begin{itemize}
\item \verb!\a! - u0007/alert/BEL
\item \verb!\b! - u0008/backspace/BB
\item \verb!\f! - u000c/form feed/FF
\item \verb!\n! - u000a/linefeed/LF
\item \verb!\r! - u000d/carriage return/CR
\item \verb!\t! - u0009/horizontal tab/HT
\item \verb!\v! - u000b/vertical tab/VT
\item \verb!\'! - u0027/single quote
\item \verb!\"! - u0022/double quote
\item \verb!\\! - u005c/backslash
\item \verb!\0! - u0000/null character/NUL 
\end{itemize}

\subsubsection{Comments}
Comments begin with the sequence \verb!/*! and end with \verb!*/!. Comments nest within each other.  Comments must be closed before the end of file is reached.

\subsubsection{Separators}
The following characters delineate various aspects of program organization (such as method arguments, array indexing, blocks, and expressions):
\begin{center}
\begin{tabular}{cccccccc}
\verb![! & \verb|]| & \verb!(! & \verb!)! & \verb!{! & \verb!,!\\
\end{tabular}
\end{center}
A notable exception is that \verb![]! itself is a lexeme related to array types and there can be no space between the two characters in this regard.